---
title: "Mapping"
author: "Gergely Som"
date: "2024-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Essentially I was wondering how to create a contour map that is iny within the extent of Budapest, but have not really managed to figure out a way to make the maps more aesthetically pleasing.

```{r} 
#Loading all the packages used in the creation of the maps
library(rJava)
library(r5r)
library(dplyr)
library(ggplot2)
library(sf)
library(osmextract)
library(osmdata)
library(interp)
library(h3jsr)
```
```{r} 
#Allocating memory for rJava
.jinit(parameters = "-Xmx8g")
```
```{r}
# 1) Build transport network
# Replace 'your_data_path' with the path to your data
your_data_path <- "C:/UCL/Dissertation/transit_data"
r5r_core <- setup_r5(data_path = your_data_path)
```

```{r}
#Reading in the base layers for the accessibility maps
budapest_districts <- st_read("Budapest_districts.kml")

budapest_outline <- st_read("C:/UCL/Dissertation/Budapest.kml")

geometries_extracted <- st_collection_extract(budapest_outline, "POLYGON")

budapest_districts$admin_level <- rep(9, nrow(budapest_districts))

# Filter out only polygon geometries
budapest_districts_poly <- budapest_districts[st_geometry_type(budapest_districts) %in% c("POLYGON", "MULTIPOLYGON"), ]
```


```{r}
#Reading in origins and destinations
orig <- st_read('orig.gpkg') |> st_transform(4326) |> st_centroid()
dest <- st_read('hospitals.gpkg') |> st_transform(4326) |> st_centroid()
```

```{r}
# For origins
orig <- orig %>%
  mutate(id = row_number(), # Create an ID column
         lat = st_coordinates(geom)[,2], # Extract latitude
         lon = st_coordinates(geom)[,1]) %>% # Extract longitude
  select(id, lat, lon) # Keep only required columns

# For destinations
dest <- dest %>%
  mutate(id = row_number(), # Create an ID column
         lat = st_coordinates(geom)[,2], # Extract latitude
         lon = st_coordinates(geom)[,1]) %>% # Extract longitude
  select(id, lat, lon) # Keep only required columns

# Add an 'opportunities' column with a placeholder value (e.g., 1 for each destination)
dest$opportunities <- 1

# Verify the structures
str(orig)
str(dest)
```


```{r}
# Running the accessibility() function for daytime accessibility
access_results <- accessibility(
  r5r_core = r5r_core,
  origins = orig,
  destinations = dest,
  opportunities_colnames = "opportunities",
  mode = "TRANSIT",
  departure_datetime = as.POSIXct("2024-03-05 12:00:00"),
  decay_function = "step",
  cutoffs = 30,
  max_trip_duration = 60,
  verbose = TRUE
)
```

```{r}
# Assuming 'orig' contains 'lat' and 'lon' columns
# Choose an appropriate resolution level (0 to 15)
res <- 8  # Example resolution level

orig_all_res <- point_to_cell(orig, res = seq(0, 15), simple = FALSE)

# Verify the H3 indexes
head(orig_all_res)
str(orig_all_res)

# retrieve polygons of H3 spatial grid
grid <- h3jsr::cell_to_polygon(orig_all_res$h3_resolution_8, simple = FALSE)

str(grid)
str(access_results)

# Add the 'id' from 'orig_all_res' to 'grid'
grid_with_id <- left_join(grid, orig_all_res[, c("h3_resolution_8", "id")], by = c("h3_address" = "h3_resolution_8"))

grid_with_id$id <- as.character(grid_with_id$id)
orig$id <- as.character(orig$id)

# merge accessibility estimates
access_sf <- left_join(grid_with_id, access_results, by = "id")
access_sf

#Clip access_sf with geometries_extracted
clipped_access_sf <- st_intersection(access_sf, geometries_extracted)

orig_df <- as.data.frame(orig)
access_hospital <- as.data.frame(clipped_access_sf)


# interpolate estimates to get spatially smooth result
access_hospitals <- access_hospital %>% 
  filter(opportunity == "opportunities") %>%
  inner_join(orig_df, by='id') %>%
  with(interp::interp(lon, lat, accessibility, duplicate = "strip")) %>%
  with(cbind(acc=as.vector(z),  # Column-major order
             x=rep(x, times=length(y)),
             y=rep(y, each=length(x)))) %>% as.data.frame() %>% na.omit() %>%
  mutate(opportunity = "opportunities")

#Inspect and save the results
str(access_hospitals)

write.csv(access_hospitals, "access_hospitals.csv", row.names = FALSE)
```

```{r}
# Assuming budapest_districts_poly is an sf object
bbox <- st_bbox(budapest_districts_poly)
bb_x <- c(bbox["xmin"], bbox["xmax"])
bb_y <- c(bbox["ymin"], bbox["ymax"])
```


```{r}
#And now doing the plotting
plot_day <- ggplot(na.omit(access_hospitals)) +
  geom_sf(data = budapest_districts_poly, fill = "lightgray", color = "black") +
  geom_contour_filled(aes(x=x, y=y, z=acc), alpha=.7, breaks = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 10, 12, 15, 18, 21, 24)) +
  scale_fill_viridis_d(direction = -1, option = 'B') +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  coord_sf(xlim = bb_x, ylim = bb_y, datum = NA) + 
  labs(fill = "Number of\nfacilities within\n30 minutes") +
  theme_void() +
  facet_wrap(~opportunity)

plot_day

```
As we can see the accessibility scores are not confined within the extend of geometries extracted which is the outline of Budapest.The question would be is there any way to fill the grey areas that the accessibility score layer does not cover and also to crop the accessibility layer with the Budapest outline.





